---
title: "Lab Week 5- Times Series and a Map"
author: "Claudia Flores"
date: "2/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

## Attach packages
```{r}
library(tidyverse)
library(janitor)
library(lubridate)
library(here)
library(paletteer)
library(tsibble)
library(fable)
library(fabletools)
library(feasts)
library(forecast)
library(sf)
library(tmap)
library(mapview)

```

## Monthly US energy consumption (renewables)
```{r}
us_renew <- read_csv(here("data","renewables_cons_prod.csv")) %>% 
  clean_names()

```

- Make description all lowercase
- Only keep observations where "description" variable contains "consumption"
- Remove any observations where "description" variable contains "total"

```{r}
renew_clean <- us_renew %>% 
  mutate(description = str_to_lower(description)) %>% 
  filter(str_detect(description, pattern = "consumption")) %>% 
  filter(!str_detect(description, pattern = "total"))

```

### Convert 'yyyymm' column to a date
ˆ
```{r}
renew_date <- renew_clean %>% 
  mutate(yr_mo_day = lubridate::parse_date_time(yyyymm, "ym")) %>% 
  mutate(month_sep = yearmonth(yr_mo_day)) %>%
  mutate(value = as.numeric(value)) %>% 
  drop_na(month_sep, value)

# Make a version where I have the month and year in separate columns
renew_parsed <- renew_date %>% 
  mutate(month = month(yr_mo_day, label = TRUE)) %>% 
  mutate(year = year(yr_mo_day))
```
### Look at it:
```{r}
renew_gg <- ggplot(data = renew_date, aes( x = month_sep, y = value, group = description)) +
  geom_line(aes(color = description)) 
  

renew_gg
```

Updating colors with paletteer palettes:
```{r}
renew_gg + 
  scale_color_paletteer_d("calecopal::figmtn")
```

### Coerce renew_parsed to a tsibble
```{r}
renew_ts <- as_tsibble(renew_parsed, key = description, index = month_sep)
```

Let's look at our time series data in a couple different ways.
```{r}
renew_ts %>% 
  autoplot(value)

renew_ts %>% 
  gg_subseries(value)

renew_ts %>% 
  gg_season(value)

ggplot(data = renew_parsed, aes(x = month, y = value, group = year)) + 
  geom_line(aes(color = year)) +
  facet_wrap(~description, ncol = 1, scales = "free", strip.position = "right")
```

### Just look at the hydroelectric energy consumption 
```{r}
hydro_ts <- renew_ts %>% 
  filter(description == "hydroelectric power consumption")

hydro_ts %>% 
  autoplot(value)

hydro_ts %>% 
  gg_subseries(value)

hydro_ts %>% 
  gg_season(value)
```

### What if I want quarterly average consumption for hydro?
```{r}
hydro_quarterly <- hydro_ts %>% 
  index_by(yearqu = ~(yearquarter(.))) %>% 
  summarize(avg_consumption = mean(value))

head(hydro_quarterly)
```

### Decompose that hydro_ts
```{r}
dcmp <- hydro_ts %>% 
  model(STL(value ~ season(window = 5)))

components(dcmp) %>% autoplot()
```
```{r}
hist(components(dcmp)$remainder)
```

Now look at the ACF:
```{r}
hydro_ts %>% 
  ACF(value) %>%
  autoplot()
```

# DANGER DANGER
```{r}
hydro_model <- hydro_ts %>% 
  model(ARIMA(value)) %>% 
  fabletools::forecast(h = "4 years")

hydro_model %>% autoplot(filter(hydro_ts, year(month_sep) > 2010))
```

### Make a world map!
```{r}
world <- read_sf(dsn = here("data","TM_WORLD_BORDERS_SIMPL-0.3-1"), 
                 layer = "TM_WORLD+BORDERS_SIMPL-0.3")

mapview(world)
```

